---
title: "Chapter 2 Group Work"
author: "Brit Riggs, Anirban Chetia, Max Evans, Helen Eifert,"
date: "9/9/2021"
output: pdf_document
---

## Question 1

### 1a. Fill in table.

```{r}
suppressPackageStartupMessages({
library(dplyr)
})

BloodType <- data.frame(
      O  = c(36, 7, 1.7, 1.5),
      A  = c(32.2, 2.9, 1.2, 0.8),
      B  = c(8.8, 2.5, 1, 0.3),
      AB = c(3.2, 0.5, 0.3, 0.1))

BloodType$Total <- BloodType$O + BloodType$A + BloodType$B + BloodType$AB

BloodType <- BloodType %>% add_row(
  O = sum(BloodType$O),
  A = sum(BloodType$A),
  B = sum(BloodType$B),
  AB = sum(BloodType$AB),
  Total = sum(BloodType$Total))

rownames(BloodType) <- c('White', 'Black', 'Asian', 'Other', 'Total')

BloodType
```

### 1b. Probability that randomly selected donor will be Asian and have Type O blood?
$P(Asian \cap Type O) = 1.7% = 0.017$

### 1c. Probability that randomly selected donor is White?
$P(White) = 80.2% = 0.802$

### 1d. Probability that randomly selected donor has Type A blood?
$P(Type A) = 37.2% = 0.372$

### 1e. Probability that a White donor will have Type A blood?
$P(Type A | White) = \frac{P(White \cap Type A)}{P(White)} = \frac{32.2}{80.2} = 0.401$

### 1f. Is blood type and ethnicity independent? Justify mathematically.
If A and B are independent, then $P(A \cap B) = P(A)P(B)$ is true.

What we know: $P(White \cap Type A) = 0.322$, $P(White) = 0.802$, and $P(Type A) = 0.372$.

Check: $P(White \cap Type A) = P(White)P(Type A) = 0.802(0.372) = 0.298 \neq 0.322$

No, blood type and ethnicity are not independent because we didn't satisfy the above equation.


## Question 2

### 2a. Number of M&Ms I eat per hour while grading homework


### 2b. The number of mornings in the coming 7 days that I change my son’s first diaper of the day.


### 2c. The number of Manzanita bushes per 100 meters of trail.


## Question 3

### 3a. What is the probability that the next two races I’m in will both have crashes?
```{r}
dbinom(2, size=2, prob=0.2)
```

### 3b. What is the probability that neither of my next two races will have a crash?
```{r}
dbinom(0, size=2, prob=0.2)
```

### 3c. What is the probability that at least one of the next two races have a crash?
```{r}
dbinom(1, size=2, prob=0.2) + dbinom(2, size=2, prob=0.2)
```


## Question 4

### 4a. What is the probability that I don’t have to clean up any vomits this coming week?
```{r}
dpois(0, lambda=1.2)
```

### 4b. What is the probability that I must clean up 1 or more vomits?
```{r}
1-dpois(0, lambda=1.2)
```

### 4c. If I wanted to measure this process with a rate per day, what rate should I use?
```{r}
(1-dpois(0, lambda=1.2))/7
```

## Question 5

### 5a. What is the probability that I see 3 runners on a morning walk?
$P(x=0) + P(x=1) + P(x=2) + P(x=3) + P(x=4+) = 1$

$0.45 + 0.25 + 0.20 + P(x=3) + 0 = 1$

$P(x=3) = 0.10$

### 5b. What is the expected number of runners that I will encounter?
Expected value, $\mu$ = (0 * (0.45)) + (1 * (0.25)) + (2 * (0.2)) + (3 * (0.1)) = 0.95 

### 5c. What is the variance of the number of runners that I will encounter?
Variance, $\sigma$ = ((0 - 0.95)^2 * 0.45)  + ((1 - 0.95)^2 * 0.25) + ((2 - 0.95)^2 * 0.2) + ((3 - 0.95)^2 * 0.1) + 0 = 1.0475

## Question 6

### 6a. $P(Z<1.58)$
```{r}
suppressPackageStartupMessages({
library(mosaic)
})
# P(Z < 1.58)
mosaic::xpnorm(1.58, 0, 1)
```

### 6b. $P(Z=1.58$
```{r}
# P(Z = 1.58) -> discrete
0
```

### 6c. $P(Z>-0.27)$
```{r}
# P(Z > -.27)
1 - mosaic::xpnorm(-.27, 0, 1)
```

### 6d. $P(-1.97<Z<2.46)$
```{r}
# Use mosaic::xpnorm for added verbose graphics/plot for visual diagnosis
# P(-1.97 < Z < 2.46)
mosaic::xpnorm(c(-1.97, 2.46), 0, 1)
```

## Question 7

### 7a. $P(Z<z)=0.75$ 
```{r}
qnorm(0.75, mean = 0, sd = 1)
```

### 7b. $P(Z>z)=0.4$
```{r}
qnorm(0.4, mean = 0, sd = 1)
```

## Question 8  

### 8a. What is the probability that I fed my cats more than 250 grams of kibble this morning?
```{r}
1 - mosaic::xpnorm(250, mean = 200, sd = 30)
```

### 8b. From my cats’ perspective, more food is better. How much would I have to feed them for this morning to be among the top 10% of feedings?


## Question 9

### 9a. Use R to find the probability of randomly sampling a sea lion with a weight greater than 320 kg. Round your answer to 3 decimals.

```{r}
round((1 - mosaic::xpnorm(320, 300, 15)), 3)
```

### 9b. Now suppose we independently sample 10 sea lions and we are interested in how many of the 10 have a weight larger than 320 kg. What distribution would we use to model this and what are the parameters of that distribution?

I would be using the Binomial Distribution as because the sea lions would either have a weight less than 320kg, or a weight more than that (either or, i.e. a binary result). The parameters for the distribution would be:
- The observations, **x**
- The number of observations (sea lions), **size**
- The probability of obtaining a sea lion with weight greater than 320kg, **p**
Function: dbinom(x, size, prob)

### 9c. Calculate by hand the probability of observing only 1 sea lion with a weight greater than 320 kg.

Calculation: We have, n = 10, and x = 1:
Formulae: \frac{n!}{x!(n-x)!}$ * 	$$\pi$^{x}$	* $(1-$\pi$)^{n-x}$
= \frac{10!}{1!(10-1)!}$ * 	$$\pi$^{1}$	* $(1-$\pi$)^{10-1}$
=

### 9d. Use R to calculate the probability of all possible outcomes and produce a graph the PMF of this distribution.

```{r echo = FALSE}
dist <- data.frame(x = 0:10) %>% 
  mutate(probability = dbinom(x, size = 10, prob = round((1 - mosaic::xpnorm(320, 300, 15)), 3)))
ggplot(dist, aes(x = x)) +
  geom_point(aes(y = probability)) +
  geom_linerange(aes(ymax = probability, ymin = 0)) +
  ggtitle('Binomial distribution: n=3, p=0.8') +
  theme_bw()
```
