---
title:  "Chapter 3 Groupwork"
author: "Brit Riggs, Anirban Chetia"
date:   "9/14/2021"
fontsize: 12pt
geometry: margin = 1.25cm
output: pdf_document
---

## Question 1

```{r, message = FALSE, warning = FALSE, echo = FALSE}
suppressPackageStartupMessages({
library(knitr)
library(dplyr)
library(boot)
library(ggplot2)
})
```

### (a)
Answer: Done (GOT IT)

```{r}
data(BodyTemp50, package = 'Lock5Data')
hist(BodyTemp50$Pulse)
```

### (b)
Answer: Done (GOT IT)

```{r}
BodyTemp50 %>% summarise(Mean = mean(Pulse), Standard.Deviation = sd(Pulse)) %>% kable()
```

### (c)
Answer: Done (GOT IT)

```{r}
mean.function <- function(x, index) {
    d <- x[index] 
    return(mean(d))  
}
Boot.BodyTemp <- boot(data = BodyTemp50$Pulse, 
                      statistic = mean.function, 
                      R = 10000)
head(Boot.BodyTemp$t)
tail(Boot.BodyTemp$t)
```

### (d)
Answer: Done (GOT IT)

```{r}
Boot.BodyTemp.graph <- data.frame(xbar = Boot.BodyTemp$t)
ggplot(Boot.BodyTemp.graph, aes(x = xbar)) +
    geom_histogram() + 
    ggtitle('Estimated Sampling distribution of the bootstrap replicates (means)')
Boot.BodyTemp.graph %>% 
  summarise(Mean.value = mean(Boot.BodyTemp$t), 
            Standard.Deviation.value = sd(Boot.BodyTemp$t)) %>% kable()
```

### (e)
Answer: Done (GOT IT)

95% confidence interval: (middle)

```{r}
quantile(Boot.BodyTemp$t, probs = c(.025, .975))
```

### (f)
Answer: Done (GOT IT)

$$
(\bar{x} - 2*\hat{\sigma}_{\bar{x}}, \bar{x} + 2*\hat{\sigma}_{\bar{x}})
$$
```{r}
paste0("(", (mean(Boot.BodyTemp$t) - 2*sd(Boot.BodyTemp$t)),
       ", ", 
       (mean(Boot.BodyTemp$t) + 2*sd(Boot.BodyTemp$t)), ")")
```
The results are almost identical to the higher end of the interval calculated in 1(e).

Could also use the *boot.ci* function:
```{r}
boot.ci(boot.out = Boot.BodyTemp, type = c("norm", "basic", "perc", "bca"))
```

## Question 2

### (a)
Answer: Done (GOT IT)

```{r}
data(EmployedACS, package = 'Lock5Data')
hist(EmployedACS$HoursWk)
```

It is not uniform and the graph appears to be right skewed a bit, as per my visual delineation.

### (b)
Answer: Done (GOT IT)

```{r}
EmployedACS %>% summarise(Mean = mean(HoursWk), Standard.Deviation = sd(HoursWk)) %>% 
  kable()
```

### (c)
Answer: Done (GOT IT)

```{r}
mean.function <- function(x, index) {
  d <- x[index]    
  return(mean(d))
}
Boot.EmployedACS <- boot(data = EmployedACS$HoursWk, 
                         statistic = mean.function, 
                         R = 10000)
head(Boot.EmployedACS$t)
tail(Boot.EmployedACS$t)
```

### (d)
Answer: Done (GOT IT)

```{r}
Boot.EmployedACS.graph <- data.frame(xbar = Boot.EmployedACS$t)
ggplot(Boot.EmployedACS.graph, aes(x = xbar)) +
  geom_histogram() + 
  ggtitle('Estimated Sampling distribution of bootstrap replicates (means)' )
Boot.EmployedACS.graph %>% 
  summarise(Mean.value = mean(Boot.EmployedACS$t), 
            Standard.Deviation.value = sd(Boot.EmployedACS$t)) %>% 
  kable()
```

### (e)
Answer: Done (GOT IT)

```{r}
quantile(Boot.EmployedACS$t, probs = c(.025, .975))
```

### (f)
Answer: Done (GOT IT)

$$
(\bar{x} - 2*\hat{\sigma}_{\bar{x}}, \bar{x} + 2*\hat{\sigma}_{\bar{x}})
$$
```{r}
paste0("(", (mean(Boot.EmployedACS$t) - 2*sd(Boot.EmployedACS$t)), 
       ", ", 
       (mean(Boot.EmployedACS$t) + 2*sd(Boot.EmployedACS$t)), ")")
```
The results are almost identical to the interval calculated in 2(e).

Again, I could also use the *boot.ci* function here:

```{r}
boot.ci(boot.out = Boot.EmployedACS, type = c("norm", "basic", "perc", "bca"))
```

## Question 3

### (a)
Answer: Done (GOT IT)

```{r}
data(BodyTemp50, package = 'Lock5Data')
standard.deviation.function <- function(x, index) {
    d <- x[index] 
    return(sd(d))  
}
```

### (b)
Answer: Done (GOT IT)

```{r}
Boot.BodyTemp <- boot(data = BodyTemp50$Pulse, 
                      statistic = standard.deviation.function, 
                      R = 10000)
```

### (c)
Answer: Done (GOT IT)

```{r}
quantile(Boot.BodyTemp$t, probs = c(.025, .975))
```
