---
title: "Chapter 4 Homework"
author: "Brit Riggs, Anirban Chetia"
date: "9/21/2021"
output: pdf_document
---

```{r, message = FALSE, warning = FALSE, echo = FALSE}
suppressPackageStartupMessages({
library(knitr)
library(dplyr)
library(boot)
library(mosaic)
library(ggplot2)
})
```

## Question 1

### (a)
Answer: Done (GOT IT)

```{r}
1 - mosaic::xpnorm(q = 230, mean = 222, sd = 5)
```

### (b)
Answer: Done (GOT IT)
$$
Expected\ value,\ E(\bar{X}) = \mu = 222
$$

### (c)
Answer: Done (GOT IT)

$$Variance,\ Var(\bar{X}) = \sigma^2_{\bar{X}} = \sigma^2/n = 25/6 = 4.1{\bar{6}}$$

### (d)
Answer: Done (GOT IT)

$$Standard Deviation,\ Std(\bar{X}) = \sqrt{\sigma^2_{\bar{X}}} = \sigma/\sqrt{n} = 5/\sqrt{6} = 2.0416$$

### (e)
Answer: Done (GOT IT)

```{r}
1 - mosaic::xpnorm(q = 230, mean = 222, sd = sqrt(25/6))
```

## Question 2

### (a)
Answer: Done (GOT IT)

```{r}
mosaic::xpnorm(q = 9, mean = 10, sd = 1.5)
```

### (b)
Answer: Done (GOT IT)

Here n = 5, therefore:
$$Standard Deviation,\ Std(\bar{X}) = \sqrt{\sigma^2_{\bar{X}}} = \sigma/\sqrt{n} = 1.5/\sqrt{5} = 0.67082039325$$

```{r}
mosaic::xpnorm(q = 9, mean = 10, sd = 1.5/sqrt(5))
```

## Question 3

### (a)
Answer: Done (GOT IT)

```{r}
mosaic::xpnorm(q = 296, mean = 298, sd = 3)
```

### (b)
Answer: Done (GOT IT)

```{r}
mosaic::xpnorm(q = 296, mean = 298, sd = 3/sqrt(6))
```

### (c)
Answer: Done (GOT IT)

```{r}
mosaic::xpnorm(q = 298, mean = 298, sd = 3)
```

### (d)
Answer: Done (GOT IT)

```{r}
mosaic::xpnorm(q = c(297, 299), mean = 298, sd = 3)
```

### (e)
Answer: Done (GOT IT)

No, the filling machine isn't out of calibration, given that the fill amount ($\mu$ = 300) is within one standard deviation (±3) of the sample mean (298).
```{r}
mosaic::xpnorm(q = 298, mean = 300, sd = 3)
```
